---
title: "Week8 Assignment07"
author: "Madison Christiansen"
date: "2023-02-03"
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE}
## Set the working directory to the root of your DSC 520 directory
setwd("/Users/madisonchristiansen/desktop")
heights_df <- read.csv("heights.csv")
```

## Linear Model 
```{r, echo=FALSE}
# Fit a linear model
earn_lm <-  lm(earn ~ height + sex + ed + age + race, data=heights_df)
# View the summary of your model
summary(earn_lm)
```

```{r, echo=FALSE}
predicted_df <- data.frame(
  earn = predict(earn_lm),
  ed=10, race=1, height=70,
  age=57, sex=1
  )
```

```{r, echo=FALSE}
## Compute deviation (i.e. residuals)
mean_earn <- mean(heights_df$earn)
## Corrected Sum of Squares Total
sst <- sum((mean_earn - predicted_df$earn)^2)
## Corrected Sum of Squares for Model
ssm <- sum((mean_earn - predicted_df$earn)^2)
## Residuals
residuals <- heights_df$earn - predicted_df$earn
## Sum of Squares for Error
sse <- sum(residuals^2)
## R Squared
r_squared <- ssm/sst
```
## Compute deviation
```{r, echo=FALSE}
mean_earn
```
## Corrected Sum of Squares Total
```{r, echo=FALSE}
sst
```
## Corrected Sum of Squares for Model
```{r, echo=FALSE}
ssm
```
## Residual
```{r, echo=FALSE}
residuals
```
## Sum of Squares for Error
```{r, echo=FALSE}
sse
```
## R Squared
```{r, echo=FALSE}
r_squared
```

```{r, echo=FALSE}
n <- 10
p <- 8
## Corrected Degrees of Freedom for Model
dfm <- p-1
## Degrees of Freedom for Error
dfe <- n-p
## Corrected Degrees of Freedom Total:   DFT = n - 1
dft <- n-1
## Mean of Squares for Model:   MSM = SSM / DFM
msm <- ssm/dfm
## Mean of Squares for Error:   MSE = SSE / DFE
mse <- sse/dfe
## Mean of Squares Total:   MST = SST / DFT
mst <- sst/dft
## F Statistic
f_score <- msm/mse
## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
adjusted_r_squared <-  1 - (1 - r_squared)*(n - 1) / (n - p)
```
## Corrected Degrees of Freedom for Model
```{r, echo=FALSE}
dfm 
```
## Degrees of Freedom for Error
```{r, echo=FALSE}
dfe
```
## Corrected Degrees of Freedom Total:   DFT = n - 1
```{r, echo=FALSE}
dft
```
## Mean of Squares for Model:   MSM = SSM / DFM
```{r, echo=FALSE}
msm 
```
## Mean of Squares for Error:   MSE = SSE / DFE
```{r, echo=FALSE}
mse
```
## Mean of Squares Total:   MST = SST / DFT
```{r, echo=FALSE}
mst
```
## F Statistic
```{r, echo=FALSE}
f_score
```
## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r, echo=FALSE}
adjusted_r_squared
```






















