---
title: "Week 8: Assignment06"
author: "Madison Christiansen"
date: "2023-02-03"
output: pdf_document
---

```{r setup, include=FALSE}
## Set the working directory to the root of your DSC 520 directory
setwd("/Users/madisonchristiansen/desktop")
heights_df <- read.csv("heights.csv")
library(ggplot2)
```

# Linear Model (Age, Earn)
```{r, echo=FALSE}
## Fit a linear model using the `age` variable as the predictor and `earn` as the outcome
age_lm <- lm(earn ~ age^2, data = heights_df)  
## View the summary of your model using `summary()`
summary(age_lm)
```

# Age Predictions
```{r}
age_predict_df <- data.frame(earn = predict(age_lm), age=20)
```

# Precictions Plotted Against Original Data
```{r, echo=FALSE}
## Plot the predictions against the original data
ggplot(data = heights_df, aes(y = earn, x = age)) +
  geom_point(color='blue') +
  geom_line(color='red',data = age_predict_df, aes(y=earn, x=age))
```

```{r, include=FALSE}
mean_earn <- mean(heights_df$earn)
## Corrected Sum of Squares Total
sst <- sum((mean_earn - heights_df$earn)^2)
## Corrected Sum of Squares for Model
ssm <- sum((mean_earn - age_predict_df$earn)^2)
## Residuals
residuals <- heights_df$earn - age_predict_df$earn
## Sum of Squares for Error
sse <- sum(residuals^2)
## R Squared R^2 = SSM\SST
r_squared <- ssm/sst
```

## Mean
```{r}
mean_earn
```

##  Squares Total
```{r}
sst
```

## Squares for Model
```{r}
ssm
```

## Residuals
```{r}
residuals
```

##Sum of Squares for Error
```{r}
sse
```

## R Squared
```{r}
r_squared
```

```{r, include=FALSE}
# number of observations
n <- 10
## Number of regression parameters
p <- 2
## Corrected Degrees of Freedom for Model (p-1)
dfm <- p -1
## Degrees of Freedom for Error (n-p)
dfe <- n-p
## Corrected Degrees of Freedom Total:   DFT = n - 1
dft <- n-1
## Mean of Squares for Model:   MSM = SSM / DFM
msm <- ssm/dfm
## Mean of Squares for Error:   MSE = SSE / DFE
mse <- sse/dfe
## Mean of Squares Total:   MST = SST / DFT
mst <- sst/dft
## F Statistic F = MSM/MSE
f_score <- msm/mse
## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
adjusted_r_squared = 1 - (1 - r_squared)*(n - 1)/(n-p)
## Calculate the p-value from the F distribution
p_value <- pf(f_score, dfm, dft, lower.tail=F)
```

## Corrected Degrees of Freedom for Model (p-1)
```{r}
dfm
```

## Degrees of Freedom for Error (n-p)
```{r}
dfe
```

## Corrected Degrees of Freedom Total:   DFT = n - 1
```{r}
dft
```

## Mean of Squares for Model:   MSM = SSM / DFM
```{r}
msm
```

## Mean of Squares for Error:   MSE = SSE / DFE
```{r}
mse
```

## Mean of Squares Total:   MST = SST / DFT
```{{r}
mst
```

## F Statistic F = MSM/MSE
```{r}
f_score
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r}
adjusted_r_squared
```

## Calculate the p-value from the F distribution
```{r}
p_value
```
















