---
title: "DSC520 Final Project (Setp3) - Plant Pollinator Networks"
author: "Madison Christiansen"
date: "2023-02-07"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

## Introduction
### This study will look at a multi-year data set consisting of data points making up the pollination systems of Missouri native plant-pollen networks from 2012 to 2020. The transfer of pollen by a pollinator is a vital step in a plant's reproductive success. The amount of pollen that can be moved by a pollinator can be impacted by the size and type of pollinator, along with the pollen's grain size. The functional group of the pollinator is looked at to categorize and analyze the ability for the pollinators to move the pollen efficiently. Pollen loads consist of heterospecific, conspecific, and corbicular loads. Looking at the specifications of the pollen and pollinators for each plant species, gives a good understanding of how the pollination systems are working, and also considerng that the climate is a huge factor as well. Drought winters are recorded in the Missouri area and this analysis can show the resiliance of plant-pollination networks when impacted by climate variations. 

## Problem Statement
### The aim of this research is to better understand the pollination systems of Missouri native plant-pollen networks to determine the impact of drought winters. Drought winters are a product of multiple things, one being climate change. Climate change is the long-term shifts in climate and weather patterns, impacting every ecosystem. Many consider climate change to be an anthropogenic issue. This research is of great interest because the pollination systems are directly linked to crops which make up most of the worlds food source. If the worlds food source is being impacted by shifts in climate this type of research is needed to help us better understand, if plant-pollen networks have reciliancy when it comes to climate changes. Understanding how these Missouri networks are impacted by drought and non-drought years then shows the larger issue of climate change and establish a small basis of understanding. 

## Adressing the Problem
### To addresss this problem I plan to analyze the data based on the different aspects of the data set. There are eight years of data collected for nine plant species, totalling around 50 smaller data sets (some years were not recorded). The data sets were then combined into a larger master dataset that I will be using for this project. The data set was collected by researchers at the Krakos Lab at Maryville University, St. Louis, MO. 
### With this being a large data set I wanted to focus on the following research questions rather then analyzing the whole.
1. Is the average pollen load consistent in size across pollinator functional groups?
2. Does the pollinator size have a impact on the amount of pollen carried?
3. Is the pollen carried conspecific or heterospecific?
4. What type of pollinators carried corbicular loads?
5. Do pollinators have consistent load sizes in drought vs non-drought years?

### These research questions will provide focus on what variables to test and how to test them. Using parametric and summary statistics I will evaluate the individual variables to then draw conclusions on the data set as a whole. A model that could be implemented would be a mathematical model and more specifically a statistical model, which is preferred for using mathematical langueage to describe the behavior of a system. 

## Analysis/Results 
```{r setup, include=FALSE}

setwd("/Users/madisonchristiansen/desktop")
plant.master <- read.csv(file="PLData.csv")
library(ggplot2)
library(readxl)
library(janitor)
library(tidyverse)
library(psych)
```

```{r,  include=FALSE}
colnames(plant.master)
new.pl.master = subset(plant.master, select = c("Plant.Species", "Pollenator.Species", "Year", "Location", "Load..", "Foreign.Pollen", "Corb.NonCorb", "Size", "Functional.Group", "Drought.Year"))
```

```{r, include=FALSE}
## deleting row 1
pl.master1 <- plant.master[-1,]
## missing variables
pl.master1[pl.master1 == ''] <- NA
pl.master2 = pl.master1 %>% drop_na()
## change string to integers , 1 == yes, no == 0
pl.master2[pl.master2== 'yes']<- 1
pl.master2[pl.master2== 'no']<- 0
```

```{r, include=FALSE}
setwd("/Users/madisonchristiansen/desktop")
Apis =  read.csv(file="Apis.csv")
bombas =  read.csv(file="bombas.csv")
Halictid =  read.csv(file="Halictid.csv")
MediumBees =  read.csv(file="MediumBees.csv")
Moth =  read.csv(file="Moth.csv")
Butterfly =  read.csv(file="Butterfly.csv")
Beetle =  read.csv(file="Beetle.csv")
Xylocopa =  read.csv(file="Xylocopa.csv")
BombBap = read.csv(file="BombBap.csv")
Apiscon = read.csv(file="Apiscon.csv")
Beetcon = read.csv(file="Beetcon.csv")
Buttcon = read.csv(file="Buttcon.csv")
Halcon = read.csv(file="Halcon.csv")
Medcon = read.csv(file="Medcon.csv")
Mothcon = read.csv(file="Mothcon.csv")
Xylcon = read.csv(file="Xylcon.csv")
Corb = read.csv(file="Corbornon.csv")
drought = read.csv(file="drought.csv")
nondrought = read.csv(file="nondrought.csv")
bombdrought = read.csv(file="bombdrought.csv")
apisdrought = read.csv(file="apisdrought.csv")
haldrought = read.csv(file="haldrought.csv")
mothdrought = read.csv(file="mothdrought.csv")
buttdrought = read.csv(file="buttdrought.csv")
xyldrought = read.csv(file="xyldrought.csv")
```

1. Is the average pollen load consistent in size across pollinator functional groups?
2. Does the pollinator size have a impact on the amount of pollen carried?

## Bombus
```{r, echo=FALSE, warning=FALSE}
describe(bombas)
```
## Apis
```{r, echo=FALSE, warning=FALSE}
describe(Apis)
```

## Halictid
```{r, echo=FALSE, warning=FALSE}
describe(Halictid)
```
## Medium Bees
```{r, echo=FALSE, warning=FALSE}
describe(MediumBees)
```
## Moth
```{r, echo=FALSE, warning=FALSE}
describe(Moth)
```
## Butterfly
```{r, echo=FALSE, warning=FALSE}
describe(Butterfly)
```
## Beetle
```{r, echo=FALSE, warning=FALSE}
describe(Beetle)
```
## Xylocopa
```{r, echo=FALSE, warning=FALSE}
describe(Xylocopa)
```
3. Is the pollen carried conspecific or heterospecific?

## Bombus
```{r, echo=FALSE, warning=FALSE}
table(BombBap['Baptisia.Fidelity'])
table(BombBap['Campanula.Fidelity'])
table(BombBap['Echinacea.paradoxa.Fidelity'])
table(BombBap['Echinacea.simulata.Fidelity'])
table(BombBap['Monarda.Fidelity'])
table(BombBap['Penstomen.cobea.Fidelity'])
table(BombBap['Penstomen.digitalis.Fidelity'])
table(BombBap['Veronicastrum.Fidelity'])
table(BombBap['Rudbeckia.Fidelity'])

```
## Apis
```{r, echo=FALSE, warning=FALSE}
table(Apiscon['Campanula.Fidelity'])
table(Apiscon['Echinacea.paradoxa.Fidelity'])
table(Apiscon['Echinacea.simulata.Fidelity'])
table(Apiscon['Monarda.Fidelity'])
table(Apiscon['Penstomen.cobea.Fidelity'])
table(Apiscon['Penstomen.digitalis.Fidelity'])
table(Apiscon['Veronicastrum.Fidelity'])
```
## Beetle
```{r, echo=FALSE, warning=FALSE}
table(Beetcon['Baptisia.Fidelity'])
table(Beetcon['Echinacea.paradoxa.Fidelity'])
table(Beetcon['Echinacea.simulata.Fidelity'])
table(Beetcon['Monarda.Fidelity'])
table(Beetcon['Rudbeckia.Fidelity'])
```
## Butterfly
```{r, echo=FALSE, warning=FALSE}
table(Buttcon['Campanula.Fidelity'])
table(Buttcon['Echinacea.paradoxa.Fidelity'])
table(Buttcon['Echinacea.simulata.Fidelity'])
table(Buttcon['Monarda.Fidelity'])
table(Buttcon['Penstomen.digitalis.Fidelity'])
```

## Halictid
```{r, echo=FALSE, warning=FALSE}
table(Halcon['Baptisia.Fidelity'])
table(Halcon['Campanula.Fidelity'])
table(Halcon['Echinacea.paradoxa.Fidelity'])
table(Halcon['Echinacea.simulata.Fidelity'])
table(Halcon['Monarda.Fidelity'])
table(Halcon['Penstomen.cobea.Fidelity'])
table(Halcon['Penstomen.digitalis.Fidelity'])
table(Halcon['Veronicastrum.Fidelity'])
table(Halcon['Rudbeckia.Fidelity'])
```

## MediumBee
```{r, echo=FALSE, warning=FALSE}
table(Medcon['Campanula.Fidelity'])
table(Medcon['Echinacea.paradoxa.Fidelity'])
table(Medcon['Echinacea.simulata.Fidelity'])
table(Medcon['Penstomen.digitalis.Fidelity'])
```

## Moth
```{r, echo=FALSE, warning=FALSE}
table(Mothcon['Campanula.Fidelity'])
table(Mothcon['Echinacea.paradoxa.Fidelity'])
table(Mothcon['Echinacea.simulata.Fidelity'])
table(Mothcon['Penstomen.digitalis.Fidelity'])
```
## Xylocopa
```{r, echo=FALSE, warning=FALSE}
table(Xylcon['Campanula.Fidelity'])
table(Xylcon['Monarda.Fidelity'])
table(Xylcon['Penstomen.digitalis.Fidelity'])
table(Xylcon['Veronicastrum.Fidelity'])
```

4. What type of pollinators carried corbicular loads?
```{r, echo=FALSE, warning=FALSE}
Corb[Corb== 'yes']<- 1
Corb[Corb== 'no']<- 0
```
## Apis
```{r, echo=FALSE, warning=FALSE}
apiscorb = subset(Corb, select = c("yes"))
apisborb1 = rename(apiscorb, Apis = yes)
table(apisborb1)
```
## Mddium Bee
```{r, echo=FALSE, warning=FALSE}
medcorb = subset(Corb, select = c("no"))
medcorb1 = rename(medcorb, MediumBee = no)
table(medcorb1)
```
## Beetle 
```{r, echo=FALSE, warning=FALSE}
betcorb = subset(Corb, select = c("no.1"))
betcorb1 = rename(betcorb, Beetle = no.1)
table(betcorb1)
```
## Bombus 
```{r, echo=FALSE, warning=FALSE}
bombcorb = subset(Corb, select = c("yes.1"))
bombcorb1 = rename(bombcorb, Bombas = yes.1)
table(bombcorb1)
```
## Butterfly 
```{r, echo=FALSE, warning=FALSE}
buttcorb = subset(Corb, select = c("no.2"))
buttcorb1 = rename(buttcorb, Butterfly = no.2)
table(buttcorb1)
```
## Halictid 
```{r, echo=FALSE, warning=FALSE}
halcorb = subset(Corb, select = c("no.4"))
halcorb1 = rename(halcorb, Halictid = no.4)
table(halcorb1)
```
## Moth 
```{r, echo=FALSE, warning=FALSE}
mothcorb = subset(Corb, select = c("no.5"))
mothcorb1 = rename(mothcorb, Moth = no.5)
table(mothcorb1)
```
## Xlyocopa 
```{r, echo=FALSE, warning=FALSE}
xlycorb = subset(Corb, select = c("yes.2"))
xlycorb1 = rename(xlycorb, Xlyocopa = yes.2)
table(xlycorb1)
```
5. Do pollinators have consistent load sizes in drought vs non-drought years?

## Bombus
```{r, echo=FALSE, warning=FALSE}
bombus2 = subset(drought, select = c("Bombus"))
describe(bombus2)
bombus3 = subset(nondrought, select = c("Bombus"))
describe(bombus3)
names(bombdrought)

bombdrought$Drought = as.numeric(as.character(bombdrought$Drought))
bombdrought$Non.Drought = as.numeric(as.integer(bombdrought$Non.Drought))
t.test(bombdrought$Drought,bombdrought$Non.Drought)
```
## Apis
```{r, echo=FALSE, warning=FALSE}
Apis2 = subset(drought, select = c("Apis"))
describe(Apis2)
Apis3 = subset(nondrought, select = c("Apis"))
describe(Apis3)

apisdrought$Drought = as.numeric(as.character(apisdrought$Drought))
apisdrought$Non.Drought = as.numeric(as.integer(apisdrought$Non.Drought))
t.test(apisdrought$Drought,apisdrought$Non.Drought)
```
## Halictid
```{r, echo=FALSE, warning=FALSE}
halictid2 = subset(drought, select = c("Halictid"))
describe(halictid2)
halictid3 = subset(nondrought, select = c("Halictid"))
describe(halictid3)

haldrought$Drought = as.numeric(as.character(haldrought$Drought))
haldrought$Non.Drought = as.numeric(as.integer(haldrought$Non.Drought))
t.test(haldrought$Drought,haldrought$Non.Drought)
```
## Medium Bee
```{r, echo=FALSE, warning=FALSE}
medbee2 = subset(drought, select = c("Medium.Bees"))
describe(medbee2)
medbee3 = subset(nondrought, select = c("Medium.Bees"))
describe(medbee3)
```
## Moth
```{r, echo=FALSE, warning=FALSE}
moth2 = subset(drought, select = c("Moth"))
describe(moth2)
moth3 = subset(nondrought, select = c("Moth"))
describe(moth3)

mothdrought$Drought = as.numeric(as.character(mothdrought$Drought))
mothdrought$Non.Drought = as.numeric(as.integer(mothdrought$Non.Drought))
t.test(mothdrought$Drought,mothdrought$Non.Drought)
```
## Butterfly
```{r, echo=FALSE, warning=FALSE}
butt2 = subset(drought, select = c("Butterfly"))
describe(butt2)
butt3 = subset(nondrought, select = c("Butterfly"))
describe(butt3)

buttdrought$Drought = as.numeric(as.character(buttdrought$Drought))
buttdrought$Non.drought = as.numeric(as.integer(buttdrought$Non.drought))
t.test(buttdrought$Drought,buttdrought$Non.Drought)
```
## Beetle 
```{r, echo=FALSE, warning=FALSE}
Beetle2 = subset(drought, select = c("Beetle"))
describe(Beetle2)
Beetle3 = subset(nondrought, select = c("Beetle"))
describe(Beetle3)
```
## Xylocopa
```{r, echo=FALSE, warning=FALSE}
Xylocopa2 = subset(drought, select = c("Xylocopa"))
describe(Xylocopa2)
Xylocopa3 = subset(nondrought, select = c("Xylocopa"))
describe(Xylocopa3)

xyldrought$Drought = as.numeric(as.character(xyldrought$Drought))
xyldrought$Non.Drought = as.numeric(as.integer(xyldrought$Non.Drought))
t.test(xyldrought$Drought,xyldrought$Non.Drought)
```
## Analysis/Results 
### To determine if the pollen loads carried by each functional group were consistent across the study species, I calculated the average of the loads collected on each pollenator, the loads are meaured by grains. You can see the average as; Bombus(large): 1,355, Apis(medium): 107, Halictid(small): 705, Medium Bees(medium): 2,015, Moth(large): 6, Butterfly(large): 258, Xylocopa(large): 3,489. Beetles did not have a sufficent sample size. To evaluate the fidelity of the pollinator functional groups I examined the composition of the pollen loads which had been designated as conspecific(0) or heterospecific(1), you can see the results above. Looking at the corbicular loads and yes(1) and no(0) give a better understand of the amount of extra pollen this pollenator functional groups carry. The pollenators with majority carrying corbicualr loads were; Apis, Bombus, Medium Bee, and Xylocopa. To determine if the pollen load amounts were consistent across pollinator functional groups in drought versus non-drought years, I compared the mean load sizes. The above results show the mean of drough as the first table and the mean of non-drought as the second table. When comparing drought to non-drought years there was no statistically difference for any functional group, Bombus (p=0.72), Apis (p=0.27), Halictidae (p= 0.96), moths (p= 0.2004), butterfly's (p= 0.90), and Xylocopa (p= 0.983). We did not have a sufficient multi-year sample size for beetles or medium bees.
## Insights
### Understanding what impacts the transfer of pollen is crucial to plant reproductive success. In keeping with previous studies, I found that certain pollinator functional groups carry different pollen load sizes. I did see high fidelity in pollen type in a pollen load across all pollinator groups based on the loads being mostly conspecific(0). Examining the corbicular loads gave insight into which pollinators are carrying older pollen, and comparing this to the other tests could give inisght as to if it imapcts thier networks. Pollen loads were consistent even with varying abiotic factors such as drought winter. This indicates that there is a resiliency to variable conditions for pollination networks. Understanding the resiliency of these networks and the limitations of this resiliency is important in predicting the impacts of climate change. Futher study is needed to undersatand these limits. 
## Limitations
### Although this study has limitations I believe my analysis provided a good understand of these Missouri pollen networks and can open the door for further studies. The limitations include gaps in the data due to collection being low some years, this caused there to be insufficient sample size for some measurments. The collection was only preformed in the summer months and does not give good repersentation of the whole year. The methods of collection may vary by year due to the often changing collectors of this data. The collection sights consisisted of two only, which could have been a large limiting factor. Improvment on this analysis would have to start by looking into the past data and collection. Since this is a eight year analysis it would be difficult to replicate all of the past data without taking another eight years. There are collectoion notebooks and methodology from past years that would be great to look at and use to futher clean up and parce the data. After understanding all of the data and starting with a clean slate, one would need to explain and track the medothology to futher build on this dataset.
## Conclusion 
### The aim of this analysis was to better understanding how Missouri networks are impacted by drought and non-drought years. This then allows us to look at the larger issue of climate change and establish a small basis of understanding. Analyzing the functional groups (pollinators) for the plant species and the pollen that is carried allows for the network to be built. Looking at all the factor surrounding the pollinators and the pollen carried gives a understanding as to if there are other extenuating circumstances that would impact these networks and allow us to rule them out. Futher giving the opportunity to focus only on the impact of drought and non-drought winters to better understand this impact exclusivly on the plant-pollination networks. This resiliency gives a good understand as to how climate change may imapct the future. 